Reconstruction of strange attractors via inter-spike intervals

### 2019-03-19 paper/main.tex
fin

Takens' theorem \cite{takens} is a delay embedding theorem, i.e it gives the conditions under which a chaotic dynamical system can be reconstructed from a sequence of observations of the state of the system. The important point is that the reconstruction preserves the properties of the system that do not change under smooth coordinate changes. Nonetheless, it does not preserve the geometric shape of structures in phase space. The theorem basically states that the strange attractor can be embedded in $k-$ dimensional Euclidean space with $k>2d_A$, being $d_A$ the box counting dimension $d_A$.

@@

Finally, as a last result, we reproduce the figure in which they change the step of predictability in order to see how \textit{far} the limit of predictability is and more important to see that there's one in fact, if there wasn't we would be doing something really wrong. We do plot in Figure 6 the predictability for different time advances where we can clearly appreciate that as we increase the number of steps ahead in time the predictability becomes smaller and smaller until a point that there's none unconditioned of the algorithm using. So yes, we can predict in some advance but it doesn't take long for chaos to strike back.

@@

\section{Conclusions}

>>>>

Takens' theorem \cite{takens} is a delay embedding theorem, i.e it gives the conditions under which a chaotic dynamical system can be reconstructed from a sequence of observations of the state of the system. The important point is that the reconstruction preserves the properties of the system that do not change under smooth coordinate changes. Nonetheless, it does not preserve the geometric shape of structures in phase space. The theorem basically states that the strange attractor can be embedded in $k-$ dimensional Euclidean space with $k>2d_A$, being $d_A$ the box counting dimension $d_A$. It is an upper constrain for the dimension needed, but we still can find our dimension to be lower than this, as we see in the Rossler system, we need to take ISI series of dimension 3.

@@

Finally, as a last result, we reproduce the figure in which they change the step of predictability in order to see how \textit{far} the limit of predictability is and more important to see that there's one in fact, if there wasn't we would be doing something really wrong. We do plot in Figure 6 the predictability for different time advances where we can clearly appreciate that as we increase the number of steps ahead in time the predictability becomes smaller and smaller until a point that there's none unconditioned of the algorithm using. So yes, we can predict in some advance but it doesn't take long for chaos to strike back. This is due to the Liapunov time, or the Liapunov exponent, that tells us when nearby trajectories will diverge thus constrining our limit of predictability

@@

\section{Conclusions}
Although confusing at first glance as the principal ideas are counter intuitive, we found here an interesting application to time series. We consider, nevertheless, that there is a lack of mathematical fundament in both papers, although the second one relies explicitly on Takens's theorem, it seems to need some general analysis. We realize how powerful this procedures explained above can be, since with little information of a dyamical system we can extract the main characteristics, reconstructing some shape of the attractor of the underlying system, establish wheter it is deterministic or stochastic, with a prediction algorithm, and even to make predictions of the future evolution, with different step-ahead. We succeeded in obtaining the main results, such that we acquired a good understanding of these topics.
